---
title: "2 Variable Inference"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(readxl)
library(modeldata)
```

## Question 1:

Null (Ho): Mean bill length is the same across species

Alternative (Ha): mean bill length is not the same across species

```{r}
#| warning: false
#| fig-alt: boxplot showing the different quartiles and averages of bill length across species.
ggplot(penguins,
       aes(x = species,
           y = bill_length_mm,
           fill = species)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2")
```

```{r}
bill_model <- aov(bill_length_mm ~ species,
                  data = penguins)

summary(bill_model)
```

Extremely statistically significant Pvalue, so we reject the null, and conclude beyond a resonable doubt that they do not have the same average bill length 

```{r}
bill_tukey <- TukeyHSD(bill_model)
bill_tukey
```

All three penguin species differ significantly in average bill length. all pvalues extremely small.

## Question 2:

Null (Ho): Sex and species are independent (i.e., each species has the same male/female proportions).

Alternative (Ha): Sex and species are not independent (i.e., proportions differ by species)

```{r}
#| fig-alt: bar chart that shows proportions of male and female birds across species. 
ggplot(penguins,
       aes(x = species,
           fill = sex,
           )) +
  geom_bar(position = "fill") +
    scale_fill_brewer(palette = "Dark2")
```

```{r}
penguin_sex <- table(penguins$sex,
                     penguins$species)

chisq.test(penguin_sex)
```

pvalue is vey large, so we can conclude that sex and species are independent of each other. 

## Question 3:

Null (Ho): Mean body mass is the same across sexes

Alternative (Ha): mean body mass is not the same across sexes

```{r}
#| fig-alt: boxplot showing quartiles and averages of body mass across bird sexes
penguins <- penguins %>% 
  filter(!is.na(body_mass_g), !is.na(sex))

#| warning: false
ggplot(penguins,
       aes(x = sex,
           y = body_mass_g,
           fill = sex)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2")
```

```{r}
sex_t_test <- t.test(body_mass_g ~ sex, 
                     data = penguins)

sex_t_test
```

pvalue is about as close to 0 as you can get, so we reject Ho and conclude that mean body mass is not the same across sexes

## Question 4

Null (Ho): Bill length and bill depth are independent

Alternative (Ha): Bill length and bill depth are not independent

```{r}
#| fig-alt: scatter plot and regresion highlighting correlations of bill length and bill depths
#| warning: false
gentoo <- penguins %>% 
  filter(species == "Gentoo", 
         !is.na(bill_length_mm),
         !is.na(bill_depth_mm))

ggplot(gentoo,
       aes(x = bill_length_mm,
           y = bill_depth_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
lm(bill_depth_mm ~ bill_length_mm,
   data = gentoo)
```

Equation is 5.1210 + .2076*(bill_length_mm)


```{r}
bill_model_2 <- lm(bill_depth_mm ~ bill_length_mm,
   data = gentoo)

summary(bill_model_2)
```

pvalue basically 0, is statistically significant, reject Ho

## Question 5:

```{r}
#| fig-alt: multiple scatter plots and regressions highliting correlation of bill length and depth across all 3 species seperately
penguins_clean <- penguins |>
  filter(!is.na(bill_length_mm), !is.na(bill_depth_mm), !is.na(species))

ggplot(penguins_clean, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_brewer(palette = "Dark2")
```

```{r}
lm_species_model <- lm(bill_depth_mm ~ bill_length_mm * species, data = penguins)
summary(lm_species_model)
```

Not exactly sure how to interpre this data. 


