---
title: "Class Notes 3/28"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(readxl)

pokemon <- read_csv("pokemon.csv")
trout <- read_excel("trout.xlsx")
```


## Git

We talked about branches a bit.

## Anova

We'll use the pokemon data set

Null: the mean HP is the same for each generation (independence)
Alternate: The mean HP is different for atleast 1 generation (dependence)

```{r}
ggplot(pokemon, aes(
  x = hp,
  y = as.factor(generation)
)) +
  geom_boxplot()
```


```{r}
pokemon <- pokemon %>% 
  mutate(generation = as.factor(generation))

aov(pokemon$hp ~ pokemon$generation)
```

to see p-value save output then get a summary

```{r}
model <- aov(hp ~ generation,
             data = pokemon)

summary(model)
```

it appears that generation does not inform HP, pvalue of .088 is too bign

#example 2 - trout

```{r}
glimpse(trout)
```


```{r}
trout <- trout %>% 
  mutate(Location = as.factor(Location))

ggplot(trout,
       aes(x = Location,
           y = Hemo)) +
  geom_boxplot() +
  geom_jitter()
```

```{r}
model <- aov(Hemo ~ Location,
             data = trout)
summary(model)
```

tiny pvalue tells us that hemo is informed by location.

(last thing today was committing to git repo)

## Class Notes 4/2



