---
title: "class notes 4.7"
format: html
embed-resources: true
---

```{r}
#| message: false

library(tidyverse)
library(broom)
```

Let's look at the `airquality` data set. First of all, are wind and temp independent?

```{r}
glimpse(airquality)
```

Null: wind and temp are independent. The slope of the regression line is zero in the population from which the sample is taken.

Alternative: they're not independent. The slope in the population is nonzero.

```{r}
#| fig-alt: a scatterplot showing the relation between wind and temp in the airquality data set. There appears to be a negative relation.

ggplot(airquality, 
       aes(x = Wind,
           y = Temp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)
```

Now let's run the test.

```{r}
model <- lm(Temp ~ Wind, 
            data = airquality)
summary(model)
```

The p-value (2.64 x 10^-9) indicates strong evidence against the null. We conclude that temp and wind are not independent. Knowing wind helps us guess temp.

## Regression from the practical perspective

Let's leave inference aside for a bit. Once we have a model, we can compute predicted (or fitted) values and also residuals. 

- fitted (predicted) values are obtained by plugging in to the regression equation.

- residuals are the difference between observed and expected values.

For observation 113, which has wind=15.5 and temp=77, the fitted value is `{r} 90.1359 - 1.2305 * 15.5`. (note my use of inline code here).

The residual of this point is (obs - exp) = `{r} 77 - (90.1359 - 1.2305 * 15.5)`. That day was almost 6 degrees warmer than the model suggested.

We can extract all fitted values and all residuals at once:

```{r}
fitted.values(model)[1:10] # first 10
residuals(model)[1:10]
```

The `broom::augment` does all this in a tidy manner.

```{r}
airq_aug <- augment(model)
glimpse(airq_aug)
```


## Class Notes 4/9: Model Diagnostics

- there is a linear relationship between the variables.
- the observations are independent of each other.
- residuals should have a normal distribution
- variance of said normal distribution should be the same for every x value

after making a regression model its pretty crucial to check these things, to make sure test is homoskedastic, among other things.






