---
title: "Class Notes 3/28"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(readxl)
library(modeldata)

pokemon <- read_csv("pokemon.csv")
trout <- read_excel("trout.xlsx")
```


## Git

We talked about branches a bit.

## Anova

We'll use the pokemon data set

Null: the mean HP is the same for each generation (independence)
Alternate: The mean HP is different for atleast 1 generation (dependence)

```{r}
ggplot(pokemon, aes(
  x = hp,
  y = as.factor(generation)
)) +
  geom_boxplot()
```


```{r}
pokemon <- pokemon %>% 
  mutate(generation = as.factor(generation))

aov(pokemon$hp ~ pokemon$generation)
```

to see p-value save output then get a summary

```{r}
model <- aov(hp ~ generation,
             data = pokemon)

summary(model)
```

it appears that generation does not inform HP, pvalue of .088 is too bign

#example 2 - trout

```{r}
glimpse(trout)
```


```{r}
trout <- trout %>% 
  mutate(Location = as.factor(Location))

ggplot(trout,
       aes(x = Location,
           y = Hemo)) +
  geom_boxplot() +
  geom_jitter()
```

```{r}
model <- aov(Hemo ~ Location,
             data = trout)
summary(model)
```

tiny pvalue tells us that hemo is informed by location.

(last thing today was committing to git repo)

## Class Notes 4/2

if the anova leads to rejecting the null, we can consider pair-wise comparisons of the groups. HOWEVER, we have to be honest about the fact that there are many possible comparisons. 

one way to do this is the *Tukey Honest Standard Differences Test*

```{r}
TukeyHSD(model)
```

## Comparing two quantitative variables:

```{r}
glimpse(crickets)
```

null: `rate` is indenpendent of `temp`

alt: `rate` is not independent of `temp`

we'll test for a linear dependence - regression

```{r}
ggplot(crickets,
       aes(x = temp,
           y = rate)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)
```

more specific:

null: the slope is zero in the population in which we are sampling

alternative: the slope is non-zero

test:

```{r}
lm(rate ~ temp,
   data = crickets) # rate explained by temp ?
```
the equation of the regression line is rate = -27.948 + 4.23(temp)

```{r}
model1 <- lm(rate ~ temp,
   data = crickets)

summary(model1)
```
pvalue is roughly 0, reject null - rate is dependent on temp.
