---
title: "Class notes 4.21"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(modeldata)
```


## Multiple regression with a categorical variable

Let's revisit the `crickets` set.

```{r}
glimpse(crickets)

ggplot(crickets, 
       aes(x = temp,
           y = rate,
           color = species)) + 
  geom_point() +
  geom_smooth(method = "lm", 
              se = FALSE) +
  scale_color_brewer(palette = "Dark2")
```

We'll assume the slopes in the two groups are equal and get two equations for the lines. For O. exclamationis, 
$rate = b_0 + b_1 * temp$ and for O. niveus, $rate = (b_0 + b_2) + b_1 * temp$.

We define the **indicator variables** `species_ON` to be 1 for O. niveus crickets and 0 otherwise. Then the model is

$$rate = \beta_0 + \beta_1 (temp) + \beta_2 (species\_ON)$$
We fit this model with R exactly as we would with quantitative variables only. 

```{r}
model <- lm(rate ~ temp + species, 
            data = crickets)
summary(model)
```

R does the coding of indicator variables automatically. The default is to look at the values of the categorical variable alphabetically. The first one is the **reference level**, the one that doesn't get an indicator variable.

$$rate = -7.21 + 3.60 (temp) - 10.07 (species\_ON)$$
We can do a residual plot here as well (but I won't).

Question: what if we don't assume the slopes are the same??

## A categorical variable with more than 2 levels

Let's look at `iris` again.

```{r}
ggplot(iris, aes(x = Sepal.Width,
                 y = Sepal.Length,
                 color = Species)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  scale_color_brewer(palette = "Dark2") 
```

Let's make a parallel-slopes model using these variables.

```{r}
model <- lm(Sepal.Length ~ Sepal.Width + Species, 
            data = iris)
summary(model)
```

The equation is

$$length = 2.2514 + 0.8036 (width) + 1.4587 (species\_vers) + 1.9468 (species\_virg)$$
For instance, a virginica with 3.0 sepal width is expected to have a length of 2.2514 + 0.8026(3.0) + 1.4587(0) + 1.9468(1) = `{r} 2.2514 + 0.8026 * 3.0 + 1.9468`.

This model corresponds to 3 equations, one for each line in the plot, one for each level of the categorical variable. For setosa, it's $length = 2.2514 + 0.8036 (width)$. For versicolor, it's $length = 2.2514 + 0.8036 (width) + 1.4587$, that is $length = 3.7101 + 0.8036 (width)$

Question: what if we want to do inference? How do we interpret these p-values?


